Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2022-07-26 Manuscript/manuscript.Rmd
update analysis

```{r initialize-packages}
library(tidyverse)
library(broom)
library(broman)
library(morris)
library(phyloseq)
library(ggpubr)
library(knitr)
library(corncob)
library(ggthemes)
library(microViz)
library(captioner)
library(knitr)
library(kableExtra)
library(flextable)
library(ggtext)
figures <- captioner()
tables <- captioner(prefix = "Table")
```

@@

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because microbiomes are often
complex, it is difficult to know a priori which microbiome members limit the
rate of an ecosystem function, and it is usually challenging to manipulate
microbiome composition directly. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response of methane oxidation rate to artificial ecosystem
selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that were enriched after selection included 24
families such as Puniceicoccaceae, Elsteraceae, and Cytophagaceae. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that artificial ecosystem selection is a useful strategy for
revealing relationships between microbiome composition and ecosystem function.

@@

Another possibility is that the microbiome itself alters the rate of ecosystem
function independent of the environment. In this case, one cannot predict the
rate of ecosystem function without understanding changes in the microbiome. In
addition, managing the microbiome through inoculation or selective antibiotics
would potentially be an effective approach for enhancing or mitigating certain
changes to ecosystem functions. These two scenarios are not mutually
exclusive, nor are they likely unique to microbes, but methods to determine the
relative important of these two scenarios have not until recently been employed.

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem of
genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
shown to respond to selection at the ecosystem-trait level [@swenson2000],
and some ecosystem-scale microbiome traits have been shown to be transferable
through microbiome inoculations [@panke-buisse2015]. Applying artificial
selection to whole microbiomes could be a novel approach for quantifying how
much of the variation in the rate of an ecosystem function could be attributed
independently to microbiome variation [@goodnight2000]. In addition, this
approach could be used to identify microbiome attributes, such as genes or taxa,
associated with the rate of an ecosystem function, although, to my knowledge
this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
Such an approach has the potential to expand our ability to model
microbial ecosystem-function relationships more accurately and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here I
use this approach to estimate how much of the variation in ecosystem function
(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
variation in the microbiome and to identify microbial markers associated with
this ecosystem function.

@@

For the current project, we applied artificial ecosystem selection on soil
microbiomes by selecting on the ecosystem-scale CH~4~ oxidation rate in order
to estimate the amount of microbiome variation underlying variation in soil CH~4~ oxidation rate
[@swenson2000]. To identify which microorganisms could be markers of CH~4~
oxidation rate, we compared the composition of the artificially selected
microbiomes to a control set of microbiomes without selection. Artificial
ecosystem selection has a similar effect to enrichment culturing by amplifying
the population of interest [@swenson2000]. This will reduce the diversity of
the soil microbiome and allow for greater power in detecting significant
markers of CH~4~ oxidation rate. We then evaluated whether these markers meet
our underlying assumptions about which taxa limit the rate of ecosystem
function.

@@

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from 0.25
g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, Inc., Waltham, MA, USA). To estimate the diversity and relative abundance of
the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
region of the 16S rRNA gene using the using the 515F - 806R primer combination
[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
$\mu$l of DNA template. Reaction conditions were: $98^\circ$C for 30 s
(initialization); 35 cycles of $98^\circ$C for 10 s (denaturation), $61^\circ$C
for 20 s (annealing), and $72^\circ$C for 20 s (extension); and $72^\circ$C for
2 m (final extension). Reactions were performed in triplicate and then combined.
Amplicons were purified twice using 0.8x ratio Mag-Bind
RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA, USA). Sequencing libraries were
prepared using a dual-indexing approach [@fadrosh2014; @kozich2013]. Amplicon
concentrations were quantified using Qubit and multiplexed at equimolar concentration. Sequencing
was performed at the University of Oregon Genomics Core Facility on the Illumina
NovaSeq 6000 with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment for the 16S reads was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' by comparing samples to negative controls [@davis2018]. We did not identify
any obvious contamination using these methods.

@@

Statistical analyses were performed in 'R' [@rcoreteam2018]. To test
whether there was a significant change in CH~4~ oxidation rate ($k$) as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. The CH~4~ oxidation rates were strongly 
right-skewed with most values close to zero and few large, positive values. This
resulted in residuals that did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed to better meet the assumptions of a linear model and to make
figures easier to interpret. First, we tested the difference of slopes between
the positive and neutral treatments by
fitting two nested models with and without the interaction between treatment
and passage using 'lm'. We compared these
models using an F-test with the 'anova' function. We
then fit the full model with the interaction term to estimate the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the mid-offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared these models using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

Richness was estimated using the method from @hurlbert1971 with a subsample size of `r min_seqs` using the 'rarefy'
function in 'vegan' [@oksanen2019]. We
tested a difference in richness between Passages 2 and 5 with both treatments
combined. Then we subset the samples from Passage 5 and compared richness between
the Positive and Neutral treatment. Next, we compared beta-diversity as the 
Bray-Curtis dissimilarity by averaging random subsets with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan' [@bray1957; @oksanen2019].
We tested a difference in centroid and dispersion of beta
diversity between the Positive and Neutral treatment and Passage 2 and 5 with
permutational analysis of variance (PERMANOVA) using the 'adonis2' function from
'vegan' [@mcardle2001; @oksanen2019].

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
assignment by grouping them at a higher level. We
then subset the samples in Passage 5 and removed all families with a prevalence
of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
pulled out the consensus taxa that were significant
with all three tests and plotted their relative abundances. For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with the Welch's t test and we used an effect size of
1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].

@@

Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across all passages
and treatments (Spearman's rho = `r myround(s[1,1], 1)`, p = `r myround(s[1,3], 1)`).

@@

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity.  Treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation. There was also an interaction
between treatment and passage, which explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45).

@@

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
*Silvanigrellaceae*, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
*Silvanigrella* is the only cultivated member of
*Silvanigrellaceae* and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

@@

Overall, none of the families enriched in the Positive selection treatment
are known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
This class is one of the groups containing methanotrophic families such as
*Methylococcaceae.* However, Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
The next family
was *Puniceicoccaceae*, which is a member of the phylum Verrucomicrobia and
contains marine microbes. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. 
An ASV from the order Armatimonadales was also enriched in the Positive treatment.
This order also contains the family *Armatimonadaceae*, which was
found to be unique to the Neutral treatment. *Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
*Gemmatimonadaceae* has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
that contains methanotrophy genes, but this family was depleted in the Positive
selection treatment [@bay2021].

# Discussion

>>>>

```{r initialize-packages}
library(tidyverse)
library(broom)
library(broman)
library(morris)
library(phyloseq)
library(ggpubr)
library(knitr)
library(corncob)
library(ggthemes)
#library(microViz)
library(captioner)
library(knitr)
library(kableExtra)
library(flextable)
library(ggtext)
figures <- captioner()
tables <- captioner(prefix = "Table")
```

@@

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because it is difficult to
manipulate microbiomes composition directly, we cannot know a priori which members of the
microbiome limit the rate of an ecosystem function, and microbiomes covary with
the environment. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that responded to selection included `r length(da_fams)`
families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that variation in microbiome composition can contribute to
variation in the rate of ecosystem function.

@@

Another possibility is that the microbiome itself alters the rate of ecosystem
function independent of the environment. In this case, one cannot predict the
rate of ecosystem function without understanding changes in the microbiome. In
addition, managing the microbiome through inoculation or selective antibiotics
would potentially be an effective approach for enhancing or mitigating certain
changes to ecosystem functions. These two scenarios are not mutually
exclusive, nor are they likely unique to microbes, but methods to determine the
relative importance of these two scenarios have not until recently been employed.

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem of
genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
shown to respond to selection at the ecosystem-trait level [@swenson2000],
and some ecosystem-scale microbiome traits have been shown to be transferable
through microbiome inoculations [@panke-buisse2015]. Applying artificial
selection to whole microbiomes could be a novel approach for quantifying how
much of the variation in the rate of an ecosystem function could be attributed
independently to microbiome variation [@goodnight2000]. In addition, this
approach could be used to identify microbiome attributes, such as genes or taxa,
associated with the rate of an ecosystem function, although, to our knowledge
this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
Such an approach has the potential to expand our ability to model
microbial ecosystem-function relationships more accurately and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here, we
use this approach to estimate how much of the variation in ecosystem function
(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
variation in the microbiome and to identify microbial markers associated with
this ecosystem function.

@@

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes without selection.  We
then evaluated whether these markers meet our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

A subsample of soil from the starting inoculum and from every jar in passages 2
and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from 0.25
g soil. Negative controls were extracted from autoclaved potting mix and
DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, Inc., Waltham, MA, USA). To estimate the diversity and relative abundance of
the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
region of the 16S rRNA gene using the 515F - 806R primer combination
[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
$\mu$l of DNA template. Reaction conditions were: $98^\circ$C for 30 s
(initialization); 35 cycles of $98^\circ$C for 10 s (denaturation), $61^\circ$C
for 20 s (annealing), and $72^\circ$C for 20 s (extension); and $72^\circ$C for
2 m (final extension). Reactions were performed in triplicate and then combined.
Amplicons were purified twice using 0.8x ratio Mag-Bind
RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA, USA). Sequencing libraries were
prepared using a dual-indexing approach [@fadrosh2014; @kozich2013]. Amplicon
concentrations were quantified using Qubit and multiplexed at equimolar concentration. Sequencing
was performed at the University of Oregon Genomics Core Facility on the Illumina
NovaSeq 6000 with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' [@davis2018].
Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance-concentration
plots indicated that `r length(contaminants)` of these were likely contaminants and these
ASVs were removed.

@@

Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
whether there was a significant change in CH~4~ oxidation rate as a
response to selection, we tested a difference in slopes between the positive
and neutral selection lines. Residuals did not meet the assumptions of constant variance
and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
transformed prior to analysis. First, we tested the difference of slopes between
the positive and neutral treatments by
fitting two nested models with and without the interaction between treatment
and passage using 'lm'. We compared these
models using an F-test with the 'anova' function. We
then fit the full model with the interaction term to estimate the slopes for each treatment, which
represented the change in CH~4~ oxidation rate per passage as a response to
selection.

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the mid-offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared nested models with and without treatment using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

Richness was estimated using the method from @hurlbert1971 with a subsample
size of `r min_seqs` using the 'rarefy' function in 'vegan' [@oksanen2019]. We
tested a difference in richness by both passage and treatment with a
Kruskal-Wallace test followed by a pairwise Wilcoxon test. Next, we estimated
beta-diversity as the Bray-Curtis dissimilarity by averaging 100 random subsets
with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan'
[@bray1957; @oksanen2019].  We tested a difference in centroid and dispersion
of beta diversity by passage and treatment using a permutational analysis of
variance (PERMANOVA) with 999 permutations using the 'adonis2' function from
'vegan' and tested a difference of group dispersions using 'betadisper' and
'anova' with 999 permutations [@mcardle2001; @oksanen2019].  Lastly, we tested
the correlation between CH~4~ oxidation rate and Bray-Curtis dissimilarity in
Passage 5 with a distance-based redundancy analysis (dbRDA) using the 'dbrda'
function in 'vegan' and estimated the p-value using a permutation F-test with
999 permutations [@mcardle2001; @oksanen2019]

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
assignment by grouping them at a higher level. We
then subset the samples in Passage 5 and removed all families with a prevalence
of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
pulled out the consensus taxa that were significant
with all three tests and plotted their relative abundances. For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we used an effect size of
1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020]. Lastly, to test differentially abundant methanotrophs, we subset all ASVs within methanotrophic families and tested their differential abundance aggregated at the family and genus level using 'corncob'.

@@

Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across the two treatments in Passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).

@@

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment with an interaction between passage and treatment (`r figures("beta", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity (F~`r tidy_permanova[1, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[1, 5], 1)`, p = `r myround(tidy_permanova[1, 6], 3)`), treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation (F~`r tidy_permanova[2, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[2, 5], 1)`, p = `r myround(tidy_permanova[2, 6], 3)`), and the interaction
between treatment and passage explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation (F~`r tidy_permanova[3, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[3, 5], 1)`, p = `r myround(tidy_permanova[3, 6], 3)`). There was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in Passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)

@@

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic assignment. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
*Silvanigrellaceae*, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
*Silvanigrella* is the only cultivated member of
*Silvanigrellaceae* and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

@@

Overall, none of the families enriched in the Positive selection treatment
contain known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
The Gammaproteobacteria include the type I and type X methanotrophs in the families *Methylococcaceae* and *Methylothermaceae* [@stein2012].
However, the Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
In addition, the *Puniceicoccaceae* is a member of the phylum Verrucomicrobia. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. Other than these two groups, none of the other taxa enriched in the Positive treatment are known to be related to methanotrophs. Two groups in the Armatimonadales were enriched in the Positive treatment including the family *Fimbriimonadaceae* and an unclassified ASV from the order Armatimonadales [@im2012].
*Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. 
The remaining families include the uncultured family 0319-6G20, *Diplorickettsiaceae*, *Rhodospirillaceae*, and an unclassified Kapabacteriales.

We did not identify any methanotrophic families in the overall differential abundance analysis. 
However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the Positive treatment.
While many *Beijerinckiaceae* are methanotrophs, several taxa in this family have lost the the ability to oxidize CH~4~ and it appears that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow on methanol [@tamas2014, @dedysh2016]. Therefore, based on this 16S analysis, it appears that no methanotrophs were enriched in the Positive treatment.

# New Discussion

We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome generates variation in the rate of ecosystem function independent of the underlying environmental conditions. While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions. To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates. We found that there was a significant increase in methane oxidation rate as a response to selection. Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation. Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem. In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars. Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.

In addition, we observed a large heritability for methane oxidation rate, which suggests that microbiome variation explains a large portion of the variation in methane oxidation rate relative to environmental variation in this system.
To estimate how much of the variation in methane oxidation rate can be explained by the presence, absence, and relative abundances of specific microbial taxa, we calculated the narrow-sense heritability of methane oxidation rate, which was `r myround(coefficients(herit_model)[4], 2)` for the Positive treatment. A heritability of 1 would indicate that a trait is entirely genetically determined and thus offspring would have the exact same value of that trait as their parents. A heritability greater than 1 indicates that offspring exceed the value of their parents. This is unlikely in the context of a genomic trait, but microbiomes are not limited by the constraints of genomes because microbes are not alleles and communities are not chromosomes. Therefore, complementarity within a microbial community could result in offspring communities that exceed the trait values of their "parent" ecosystems. The large heritability we observed indicates that microbiome variation is a large contributor to the variation in methane oxidation in this experimental system.

Notably, the Neutral selection treatment displayed a negative heritability of `r myround(coefficients(herit_model)[2], 2)`. Negative heritability estimates can arise in artificial selection experiments. One explanation for this phenomenon is that traits that are favored under the experimental conditions are negatively correlated with the trait under selection. In this experiment, CH~4~ oxidation rate could be negatively correlated with traits that favor persistence in jars [@conner2003]. This tradeoff would result in a negative heritability because, in the absence of an imposed selection pressure, passaging in jars would favor microbes with lower CH4 oxidation rates.

Next, we wanted to address what aspects of the microbiome, in terms of alpha diversity, beta diversity, or individual taxa, explain this contribution of the microbiome to variation in methane oxidation rate. There are three ways that the microbiome could respond to selection in this experiment: gain or loss of species, changes in the relative abundances of species, or changes within the genomes of the constituent species. Using 16S rRNA gene sequences, we are unable to address the third possibility of whether taxa in this experiment could have evolved changes in their genomes as a result of selection. Therefore, we will focus on the first two possibilities.

Richness at the ASV level did not vary between the two treatments and there were relatively few taxa gained or lost in the Positive selection treatment and none of these were prevalent across the 12 jars in passage 5. Therefore, the gain or loss of species is unlikely to explain the increase in methane oxidation rate. However, we found that Bray-Curtis dissimilarity was greater between the two treatments in Passage 5 than within each treatment and was correlated with methane oxidation rate, which suggests that changes in the relative abundance of taxa could explain the response to selection. A difference in composition between the two treatments could be a direct result of selection for organisms that contribute to CH~4~ oxidation or a result of ecological drift over the course of the experiment. In the former case, that would suggest that changes in the relative abundance of specific microbial taxa could explain the response to selection on CH~4~ oxidation rate.

Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs. This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?). In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite. I know that one methanogenesis paper, not sure about methanotrophy). However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed in increase in methane oxidation rate, we may have to look at other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.

One of the goals of this study was to determine how much of the variation in ecosystem function is due to variation in the microbiome. In this experiment, we eliminated as much environmental variation as possible. This was intentional, because we wanted to determine whether it was possible for variation in the microbiome to generate variation in methane oxidation rate independent of the environment. As a result, our heritability estimates are likely to be greater than they would be in natural ecosystems with greater environmental variation. Future work could estimate heritability in experimental systems that manipulate environmental variation or across natural ecosystems to evaluate how much of the variation in ecosystem function can be explained by microbiome variation outside the lab. This experiment also had a relatively small sample size which resulted in heritability estimates with wide error bars. Increasing replication would help us more accurately estimate this parameter and give us a better understanding of the relative importance of environmental variation and microbiome variation in the rate of microbiome-mediated ecosystem functions.

Another important question is whether selection at the ecosystem scale resulted in within-strain evolution. Using 16S rRNA gene sequencing we are unable to resolve genomic changes within taxa that may have undergone evolution. Future work could analyze an artificial ecosystem selection experiment using genome-resolved metagenomics to try to detect strain-level changes in the constituent taxa. This would help evaluate whether certain metabolic pathways were enriched in the positive selection treatment or if functional genes changed in frequency within the microbiome.

As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function. Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function. However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously using machine learning methods [@topçuoğlu2020]
